
---
title: "Modeling and Representing Data HW3"
author: "Derek Wales"
date: "27SEP19"
output:
  pdf_document: default
  html_document:
    highlight: pygments
    theme: spacelab
---

```{r setup, include=FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
# The very first time you use this R markdown file, you should install each of the packages below.
# The same goes for other packages you might decide to use
# Remember that you only need to install each R package once in R (ever -- unless you change computers).
# All you need to do whenever you need to use the package again (after restarting the R session),
# is to use the library function to call the package.
# For example, type install.packages("knitr") in the console to install the knitr package. 
knitr::opts_chunk$set(echo = TRUE)
library(lme4)
library(tidyverse)
library(ggplot2)
library(ggpubr)
library(MASS)
```

* * *
```{r echo=FALSE}
#Reading file with data for all the voters
all_voter = read.csv("C:\\Users\\derek\\Desktop\\MIDS 1st Semester\\Modeling and Representing Data\\Team Projects\\Project Two\\team-project-2-estrogen-bioassay-and-voting-in-nc-avengers\\Data\\voter_stats_20161108.txt", 
                     header = TRUE, comment.char = "",sep='', stringsAsFactors = F)
#All voters -  514,846 Observations  (Actual number of rows in file)

#Getting unique rows from all_voters
all_voter = unique(all_voter)

#All voters -  461,833 Observations

#Reading file with data for the voters who actually voted 
voted_voter = read.delim("C:\\Users\\derek\\Desktop\\MIDS 1st Semester\\Modeling and Representing Data\\Team Projects\\Project Two\\team-project-2-estrogen-bioassay-and-voting-in-nc-avengers\\Data\\history_stats_20161108.txt")

#Voted voters - 734,126 Observations    (Actual number of rows in file)
#all unique rows in Voted_voters already

#renaming total_voter column in Voted_voters dataframe
colnames(voted_voter)[9] <- "voted_voters" # I recommend changing to voters and total voters to 
                                           # Registered Voters

#Changing datatype of votes to numeric in both the dataframes
all_voter$total_voters = as.numeric(all_voter$total_voters)
voted_voter$voted_voters = as.numeric(as.character(voted_voter$voted_voters))

#Calculating total voting percentage 
Num_all_voters <- sum(all_voter$total_voters, na.rm = T)  #6,213,883  (Total voters)
Num_of_voted_voters <- sum(voted_voter$voted_voters)      #4,768,079  (Voters who voted)
Percent_Voted = (Num_of_voted_voters*100)/Num_all_voters  #76.73 % Voters Voted

#Removing unwanted columns (DATES & constant value stats_type columns)
all_voter$election_date <- NULL  
voted_voter$election_date <- NULL
voted_voter$update_date <- NULL
all_voter$stats_type <- NULL
voted_voter$stats_type <- NULL

#Removing Rows with empty data
#inital count of rows |  all_voter - 461,833, voted_voter - 734,126

all_v_noNA <- all_voter %>%
  na_if("") %>%
  na.omit()
#deleted 1026 rows with missing values (0.2%)

voted_v_noNA <- voted_voter %>%
  na_if("") %>%
  na.omit()
#deleted 30,024 rows with missing values (4.0%)

#new count of rows  |   all_voter - 460,807, voted_voter  - 704,102

Num_all_voters <- sum(all_v_noNA$total_voters, na.rm = T)  #6,210,364  (Total voters)
Num_of_voted_voters <- sum(voted_v_noNA$voted_voters)      #4,572,359  (Voters who voted)
Percent_Voted = (Num_of_voted_voters*100)/Num_all_voters   #73.62 % Voters Voted

#merging the two dataframes 
votedDataMerged = voted_v_noNA %>% 
  inner_join (all_v_noNA, 
              by = c("county_desc" , "precinct_abbrv", "vtd_abbrv", 
                     "party_cd", "race_code","ethnic_code","sex_code",       
                     "age"))
#nrows in votedDataMerged 626,544

#removing noNA dataframes (not required anymore)
rm(voted_v_noNA,all_v_noNA)

#dropping voting_method column (same as voting_method_desc)
votedDataMerged$voting_method <- NULL

# What's this?
str(votedDataMerged)
votedDataMerged$county_desc = as.factor(votedDataMerged$county_desc)
votedDataMerged$precinct_abbrv = as.factor(votedDataMerged$precinct_abbrv)
votedDataMerged$vtd_abbrv = as.factor(votedDataMerged$vtd_abbrv)
votedDataMerged$age = as.factor(votedDataMerged$age)
votedDataMerged$party_cd = as.factor(votedDataMerged$party_cd)
votedDataMerged$race_code = as.factor(votedDataMerged$race_code)
votedDataMerged$ethnic_code = as.factor(votedDataMerged$ethnic_code)
votedDataMerged$sex_code = as.factor(votedDataMerged$sex_code)

# changing already factored variable to character and then again to factor (removing any extra factor)
votedDataMerged$voting_method_desc = as.character(votedDataMerged$voting_method_desc)
votedDataMerged$voting_method_desc = as.factor(votedDataMerged$voting_method_desc)
votedDataMerged$voted_party_cd = as.character(votedDataMerged$voted_party_cd)
votedDataMerged$voted_party_cd = as.factor(votedDataMerged$voted_party_cd)

nrow(unique(votedDataMerged[,c("county_desc" , "precinct_abbrv", "vtd_abbrv",
                               "party_cd", "race_code","ethnic_code","sex_code", 
                               "age","voted_voters","voting_method_desc","voted_party_cd")]))

# Creating a new dataframe to get Overall Voting stats 
# Since the same number of total_voter was appearing for different voted_voters in votedDataMerged
# (which differed by voting method and voted_party code), a new dataframe has to be created 
# (Akshay explained) this to me.

voterStatDf <- aggregate(votedDataMerged$voted_voters, 
                         by=list(votedDataMerged$county_desc , votedDataMerged$precinct_abbrv, 
                                 votedDataMerged$vtd_abbrv,votedDataMerged$party_cd, votedDataMerged$age,
                                 votedDataMerged$race_code,votedDataMerged$ethnic_code,votedDataMerged$sex_code,
                                 votedDataMerged$total_voters), sum) 

colnames(voterStatDf) <- c("county_desc", "precinct_abbrv", "vtd_abbrv" ,"party_cd",   
                           "age", "race_code", "ethnic_code", "sex_code",  "total_voters",        
                           "voted_voters")

#Calculating total voting percentage after merging
Num_all_voters <- sum(voterStatDf$total_voters)             #6,070,763 (Total voters)
Num_of_voted_voters <- sum(voterStatDf$voted_voters)        #4,097,895  (Voters who voted)
Percent_Voted = (Num_of_voted_voters*100)/Num_all_voters    #67.52% Voters Voted

#Checking number & percentage of observations which has more Voted votes than total voters
## in Voter stats dataframe (Anomaly more people voted than registered voters)
nrow(voterStatDf[voterStatDf$voted_voters > voterStatDf$total_voters,  ]) # 3023 (0.84%)

## in voter data merged dataframe
nrow(votedDataMerged[votedDataMerged$voted_voters > votedDataMerged$total_voters,  ]) # 622 (0.1%)

#counting and plotting number of observations for each county
dt = aggregate(voterStatDf$county_desc, list(voterStatDf$county_desc), length)
colnames(dt) = c('County','NbrOfObs')
```

### Summary:

For this project our team was asked to look at voter registration, and participation across the state of North Carolina. To do this we used a series of tools that had been taught in class, including critical thinking, exploratory data analysis, logistic regression, outside research, and hierarchical modeling techniques. 

Our analysis was targeted at answers the following research questions.

•	How did demographic subgroups vote in 2016? For example, how did the turnout for males compare to the turnout for females after controlling for other potential predictors?
•	Did the overall probability or odds of voting differ by county in 2016?
•	How did the turnout rates differ between females and males for the different party affiliations?

Our analysis ultimately led to the following model:

#### 




####

### Intro:

The North Carolina State Board of Elections (NCSBE) is administration and the election process for the State. They are also required to keep extensive records to ensure electoral compliance, as part of their duties, they also keep information on likely voters and registered voters. 

For this project we were given a dataset that contained information about the 2016 election, and we were asked to use two separate datasets, one which contained registered voters and the other which had likely voters. 

From that we needed to clean, import, explore, and ultimately model these two datasets in R to answer the above research questions. 

### Data and Exploratory Data Analysis:

One of the more challenging aspects of this project was getting the two separate dataframes (registered voters and voters) merged into one dataframe that could be analyzed. Initial attempts to merge resulted in R crashing because of missing values, additionally my first successful join deleted all of the registered voters data because it had the same column name as the total voters. 

In order to successfully merge the two dataframses the team went through a series of steps to ensure a successful merge. This was first done by removing the null values in both dataframes, which still left us with over a million observations between the two dataframes. To validate the success of this operation, we then took the total voters over the registered voters and confirmed it matched roughly matched the voter turnout percentage in 2016 (which it did). 

We were then able to merge the two dataframes and begin working towards what we would use for our final analysis (named VotedDataMerged). From here we could check and ensure all variables were the right type (for example county names and race as factors instead of strings).

From this we could then aggregate into the VoterStatDF. The team then validated that we did not have any unusual values/percentages, things like the ensuring the amount of voters exceeded (which only happened in 0.84% of cases, or that the voter turnout was close to the average across the state, etc).  
From here, the team began graphically exploring variables within the dataset this starting with the total voters, then exploring the voter participation rate within each (voters vs registered voters). After that we selected a random sample of 20 different counties (varying in size and  demographics) across the state. 

With that sample we then began graphing how much the various ethic and gender groups participated. First looking at gender, then each of the ethnic groups. From our initial plots we could see that voting in the US was very split along ethnic, gender, age and party lines. 

Additionally, we also did multiple plots that overlaid various factors (age and race etc). From this we could see that each group sub group had influence on voting patterns. Because of this we started modeling with a Hierarchal Model. 
 
### Model Selection Process and Assessment:



### Final Model

\[\log(\frac{\pi_i}{1-\pi_i}) = \beta_0 + \beta_1 Black + \beta_2 Asian - \beta_3 MotherWeight +  \beta_3 Smoking\]
where $y_i|x_i \sim$ Bernouilli($\pi_i$)

\newpage 
### Coefficients in table below

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
(Intercept) & -0.4634 & 0.6149 & -0.75 & 0.4511 \\ 
  Black & 0.7576 & 0.2116 & 3.58 & 0.0003 \\ 
  Asain & 0.8124 & 0.3954 & 2.05 & 0.0399 \\ 
  Mother Weight & -0.0109 & 0.0047 & -2.31 & 0.0211 \\ 
  Smoker & 0.3614 & 0.1790 & 2.02 & 0.0435 \\ 
   \hline
\end{tabular}
\end{table}

### Conclusion/Remarks:

•	How did demographic subgroups vote in 2016? For example, how did the turnout for males compare to the turnout for females after controlling for other potential predictors?

•	Did the overall probability or odds of voting differ by county in 2016?

•	How did the turnout rates differ between females and males for the different party affiliations?

\newpage

## Appendix (All EDA and Model Building)

```{r}
ggplot(dt)+
  geom_bar(aes(x=County, y=NbrOfObs), stat = 'identity')+
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1, size = 3))


#counting and plotting number of total voters and voted voters for each county
dt = aggregate(voterStatDf$total_voters, list(voterStatDf$county_desc), sum)
colnames(dt) = c('County','NbrOfTotalVoters')
dt2 = aggregate(voterStatDf$voted_voters, list(voterStatDf$county_desc), sum)
colnames(dt2) = c('County','NbrOfVotedVoters')
dt3 = merge(dt, dt2, by = "County")
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))

ggplot(dt3, aes(x=County))+
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.4)+
  ggtitle("Distribution of Total Voter and Voted voters for 20 Counties") +
  geom_text(data=dt3,aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            vjust = 0.6, hjust= -0.2, angle=90, size=2.5) +
  xlab("Counties")+
  ylab("Voter Count") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1,size=2))

#############################################################################################

#Selecting 20 counties at rondom

set.seed(98)
counties = sample(as.character(unique(votedDataMerged$county_desc)), size = 20,replace = T)

#Couting number of observations of 20 selected counties 

print(counties)

# [1] "BERTIE"     "CALDWELL"   "FRANKLIN"   "DARE"      
# [5] "HARNETT"    "JONES"      "CHATHAM"    "SWAIN"     
# [9] "PAMLICO"    "BUNCOMBE"   "CHEROKEE"   "MCDOWELL"  
# [13] "STOKES"     "PITT"       "NASH"       "SURRY"     
# [17] "WARREN"     "TYRRELL"    "LENOIR"     "CUMBERLAND"

#colnames(voting_dataset) - 12 columns
# county_desc   | 
# precinct_abbrv| vtd_abbrv | 
# race_code     | sex_code  | age | ethnic_code
# total_voters  |  party_cd  
# voting_method_desc | voted_voters | voted_party_cd

#Subsetting data for 20 counties from merged and stats table

voting_dataset = subset(votedDataMerged, county_desc %in% counties)
voting_stats_dataset = subset(voterStatDf, county_desc %in% counties)

#Calculating total voting percentage after sampling (makes sense)

Num_all_voters <- sum(voting_stats_dataset$total_voters)        #947,012 (Total voters)
Num_of_voted_voters <- sum(voting_stats_dataset$voted_voters)   #648,649  (Voters who voted)
Percent_Voted = (Num_of_voted_voters*100)/Num_all_voters        #68.49% Voters Voted


#Counting and Plotting the number of observations for 20 couties 

dt = aggregate(voting_stats_dataset$county_desc, list(voting_stats_dataset$county_desc), length)
colnames(dt) = c('County','NbrOfObs')

ggplot(dt,aes(x=County, y=NbrOfObs),label=NbrOfObs)+
  geom_bar(stat='identity',fill='Blue2', width = 0.4)+
  ylim(0,15500)+
  geom_text(aes(label= scales::comma(NbrOfObs)),hjust=0.5,vjust=-0.8,size=3)+
  ggtitle("Number of Observations for 20 Counties") +
  theme( plot.title = element_text(hjust = 0.5),
         axis.text.x = element_text(angle = 45,hjust = 1))


#Checking number observations where party has changed
nrow(voting_dataset[voting_dataset$party_cd != voting_dataset$voted_party_cd,]) #868 (0.8%)

# Since only 0.8% observations have different partycd and voted_party_cd. 
# Dropping party_cd column

voting_dataset$party_cd <- NULL

# Plotting some EDA

#############################################################################################

## Calculating County-wise total and voted voters per county along with %voted

dt = aggregate(c(voting_stats_dataset$total_voters), list(voting_stats_dataset$county_desc), sum)
colnames(dt) = c('County','NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), list(voting_stats_dataset$county_desc), sum)
colnames(dt2) = c('County','NbrOfVotedVoters')
dt3 = merge(dt, dt2, by = "County")
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))

## Plotting number of total voters and count of voters who voted

ggplot(dt3, aes(x=County))+
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.4)+
  ggtitle("Distribution of Total Voter and Voted voters for 20 Counties") +
  geom_text(data=dt3,aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            vjust = 0.6, hjust= -0.2, angle=90, size=2.5 ) +
  ylim(0,200000)+
  xlab("Counties")+
  ylab("Voter Count") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))

#############################################################################################

# Gender wise voter turnout

## Calculating Gender-wise total and voted voters per county along with %voted

dt = aggregate(c(voting_stats_dataset$total_voters), 
               list(voting_stats_dataset$sex_code), sum)
colnames(dt) = c('sex_code','NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), 
                list(voting_stats_dataset$sex_code), sum)
colnames(dt2) = c('sex_code','NbrOfVotedVoters')
dt3 = merge(dt, dt2, by = "sex_code")
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))

## Plotting number of total voters and count of voters who voted

ggplot(dt3, aes(x=sex_code))+
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Gender-wise Voter Turnout") +
  geom_text(data=dt3,aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            vjust = -0.4, hjust= 0.4, angle=0, size=3 ) +
  xlab("Gender")+
  ylab("Voter Count") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))

#############################################################################################

## Plotting Voter turnout based on county and gender

# Calculating total and voted voters per county along with %voted
dt = aggregate(c(voting_stats_dataset$total_voters), 
               list(voting_stats_dataset$county_desc,voting_stats_dataset$sex_code), sum)
colnames(dt) = c('County',"Sex_code",'NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), 
                list(voting_stats_dataset$county_desc,voting_stats_dataset$sex_code), sum)
colnames(dt2) = c('County',"Sex_code", "NbrOfVotedVoters")
dt3 = merge(dt, dt2, by.x = c("County","Sex_code"), by.y=c("County","Sex_code"))
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))

## plotting number of total voters and count of voters who voted 

## County Wise

ggplot(dt3, aes(x=County))+
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2')+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2')+
  ggtitle("County-wise Voter Turnout") +
  geom_text(data=dt3,aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = -0.1, vjust = 0.5, angle=90, size=3) +
  ylim(0,105000)+
  xlab("Counties")+
  ylab("Voter Count") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))+
  facet_wrap(~Sex_code)

## Gender Wise 

ggplot(dt3, aes(x=Sex_code))+
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.6)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.4)+
  ggtitle("Gender-wise voter Turnout") +
  geom_text(data=dt3,aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  ylim(0,105000)+
  xlab("Counties")+
  ylab("Voter Count") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))+
  facet_wrap(~County)

### Based on the analysis, it seems for each county, the number of males and females who voted
### were comparable (+/- 2%)

#############################################################################################

#Calculating and plotting Party wise voter turnout according to Gender

## Calculating. Making dataframe for new values and groups 

dt = aggregate(c(voting_stats_dataset$total_voters), 
               list(voting_stats_dataset$party_cd,voting_stats_dataset$sex_code), sum)
colnames(dt) = c('party_cd',"Sex_code",'NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), 
                list(voting_stats_dataset$party_cd,voting_stats_dataset$sex_code), sum)
colnames(dt2) = c('party_cd',"Sex_code", "NbrOfVotedVoters")
dt3 = merge(dt, dt2, by = c("party_cd","Sex_code"))
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))

## Plotting

female_party_plot <- ggplot(dt3[dt3$Sex_code=="F",], aes(x=party_cd))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Party-wise Voter-Turnout (Female)") +
  geom_text(data=dt3[dt3$Sex_code=="F",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Party Code")+
  ylab("Female Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))

male_party_plot <- ggplot(dt3[dt3$Sex_code=="M",], aes(x=party_cd))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Party-wise Voter-Turnout (Male)") +
  geom_text(data=dt3[dt3$Sex_code=="M",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Party Code")+
  ylab("Male Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))


ggarrange(male_party_plot,female_party_plot, ncol = 2)

#############################################################################################

#Calculating and plotting Age wise voter turnout according to Gender

dt = aggregate(c(voting_stats_dataset$total_voters), 
               list(voting_stats_dataset$age,voting_stats_dataset$sex_code), sum)
colnames(dt) = c('age',"Sex_code",'NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), 
                list(voting_stats_dataset$age,voting_stats_dataset$sex_code), sum)
colnames(dt2) = c('age',"Sex_code", "NbrOfVotedVoters")
dt3 = merge(dt, dt2, by = c("age","Sex_code"))
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))


female_age_plot <- ggplot(dt3[dt3$Sex_code=="F",], aes(x=age))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Age wise Voter-Turnout (Female)") +
  geom_text(data=dt3[dt3$Sex_code=="F",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Age Groups")+
  ylab("Female Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))

male_age_plot <- ggplot(dt3[dt3$Sex_code=="M",], aes(x=age))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Age wise Voter-Turnout (Male)") +
  geom_text(data=dt3[dt3$Sex_code=="M",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Age Groups")+
  ylab("Male Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))


ggarrange(female_age_plot,male_age_plot, ncol = 2)

#############################################################################################

#Calculating and plotting Race wise voter turnout according to Gender

dt = aggregate(c(voting_stats_dataset$total_voters), 
               list(voting_stats_dataset$race,voting_stats_dataset$sex_code), sum)
colnames(dt) = c('race',"Sex_code",'NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), 
                list(voting_stats_dataset$race,voting_stats_dataset$sex_code), sum)
colnames(dt2) = c('race',"Sex_code", "NbrOfVotedVoters")
dt3 = merge(dt, dt2, by = c("race","Sex_code"))
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))


female_race_plot <- ggplot(dt3[dt3$Sex_code=="F",], aes(x=race))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Race wise Voter-Turnout (Female)") +
  geom_text(data=dt3[dt3$Sex_code=="F",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Race")+
  ylab("Female Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))

male_race_plot <- ggplot(dt3[dt3$Sex_code=="M",], aes(x=race))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Race wise Voter-Turnout (Male)") +
  geom_text(data=dt3[dt3$Sex_code=="M",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Race")+
  ylab("Male Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))


ggarrange(female_race_plot,male_race_plot, ncol = 2)

#############################################################################################

#Calculating and plotting Race wise voter turnout according to Ethinic Group

dt = aggregate(c(voting_stats_dataset$total_voters), 
               list(voting_stats_dataset$ethnic_code,voting_stats_dataset$sex_code), sum)
colnames(dt) = c('ethnic_code',"Sex_code",'NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), 
                list(voting_stats_dataset$ethnic_code,voting_stats_dataset$sex_code), sum)
colnames(dt2) = c('ethnic_code',"Sex_code", "NbrOfVotedVoters")
dt3 = merge(dt, dt2, by = c("ethnic_code","Sex_code"))
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))


female_ethnic_plot <- ggplot(dt3[dt3$Sex_code=="F",], aes(x=ethnic_code))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Ethnic-code wise Voter-Turnout (Female)") +
  geom_text(data=dt3[dt3$Sex_code=="F",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Ethnic-code")+
  ylab("Female Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))

male_ethnic_plot <- ggplot(dt3[dt3$Sex_code=="M",], aes(x=ethnic_code))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Ethnic-code wise Voter-Turnout (Male)") +
  geom_text(data=dt3[dt3$Sex_code=="M",],
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Ethnic-code")+
  ylab("Male Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))


ggarrange(female_ethnic_plot,male_ethnic_plot, ncol = 2)

#############################################################################################

#Calculating and plotting Age-wise wise voter turnout according to Ethinic Group

dt = aggregate(c(voting_stats_dataset$total_voters), 
               list(voting_stats_dataset$ethnic_code,voting_stats_dataset$age), sum)
colnames(dt) = c('ethnic_code',"age",'NbrOfTotalVoters')
dt2 = aggregate(c(voting_stats_dataset$voted_voters), 
                list(voting_stats_dataset$ethnic_code,voting_stats_dataset$age), sum)
colnames(dt2) = c('ethnic_code',"age", "NbrOfVotedVoters")
dt3 = merge(dt, dt2, by = c("ethnic_code","age"))
dt3 = mutate(dt3, percentVoted = round((NbrOfVotedVoters/NbrOfTotalVoters),2))


agewise_ethnic_plot <- ggplot(dt3, aes(x=age))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Age-wise Voter-Turnout for Ethnic groups") +
  geom_text(data=dt3,
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Age")+
  ylab("Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))+
  facet_wrap(~ethnic_code)

ethnicwise_age_plot <- ggplot(dt3, aes(x=ethnic_code))+ 
  geom_bar(aes(y=NbrOfTotalVoters), stat = 'identity', fill='Blue2', width = 0.4)+
  geom_bar(aes(y=NbrOfVotedVoters), stat = 'identity', fill='Orange2', width = 0.2)+
  ggtitle("Ethnic group wise Voter-Turnout for Age groups") +
  geom_text(data=dt3,
            aes(y=NbrOfTotalVoters,label=scales::percent(percentVoted)), 
            hjust = 0.5, vjust = -0.5, angle=0, size=3) +
  xlab("Ehnic Code")+
  ylab("Voters") +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = 45,hjust = 1))+
  facet_wrap(~age)

ggarrange(agewise_ethnic_plot,ethnicwise_age_plot, ncol = 2)

#############################################################################################

# EDA Complete Moving into Modeling
# Use voting_stats_dataset
colnames(voting_stats_dataset)

# [1] "county_desc"    "precinct_abbrv" "vtd_abbrv"      "party_cd"       "age"           
# [6] "race_code"      "ethnic_code"    "sex_code"       "total_voters"   "voted_voters" 

# All factors included, with varying effects by county. 
model_full <- glmer(cbind(total_voters, voted_voters) ~ (1|county_desc) + 
                 race_code + ethnic_code + age + party_cd + sex_code, 
               data = voting_stats_dataset, family = binomial)

summary(model_full)

```



* * *



