---
title: "Estrogen Bioassay"
output:
  pdf_document: default
  html_document: default
---

```{r s, inlcude = FALSE}
library(ggplot2)
library(lme4)
library(readr)
library(arm)
library(car)
library(e1071)
library(caret)
library(pROC)
```

```{r setup, include = FALSE}

estrogen <- read.table('~/Desktop/bioassay.txt', header=T, sep=" ", colClasses = c('protocol' = 'factor'))

#convert factors to numeric
estrogen$uterus <- as.character(estrogen$uterus)
estrogen$uterus <- as.numeric(estrogen$uterus)
estrogen$weight <- as.character(estrogen$weight)
estrogen$weight <- as.numeric(estrogen$weight)
estrogen$group <- as.factor(estrogen$group)
#remove empty data
estrogen <- na.omit(estrogen)
#need to mean center weight
estrogen$mweight <- estrogen$weight - mean(estrogen$weight)

# added another binary variable to distinguish between mature and inmature rats
estrogen$mature <- 1
estrogen$mature[estrogen$protocol == 'A' | estrogen$protocol == 'B' ] <- 0
estrogen$mature <- as.factor(estrogen$mature)

summary(estrogen)
```


```{r EDA, include = FALSE}
# check the distribution of the response variable
hist(estrogen$uterus)
hist(log(estrogen$uterus))
#Log looks the best compared to the other transformations, use this for analysis
hist(sqrt(estrogen$uterus))
hist((estrogen$uterus)^(1/3))
```


```{r EDA2, include = FALSE}

ggplot(data = estrogen, aes(x=mweight, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3') + facet_wrap(~protocol, ncol = 3)
#the mature rats have a negative pattern, the immature rats do not appear to have a pattern

ggplot(data = estrogen, aes(x=EE, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3') + facet_wrap(~protocol, ncol = 3)
#There appears to be a positive relationship between EE and log uterus weight

ggplot(data = estrogen, aes(x=ZM, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3') + facet_wrap(~protocol, ncol = 3)
#There appears to be a negative relationship between ZM and uterus weight

ggplot(data = estrogen, aes(x=lab, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3') + facet_wrap(~protocol, ncol = 3) + theme(axis.text.x = element_text(angle = 90))
#It appears that labs doing similar experiments have similar outcomes

ggplot(data = estrogen, aes(x=lab, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3') + theme(axis.text.x = element_text(angle = 90))
#See differences in labs, but this could be due to some do not do certain protocols. Need to build the model to check

ggplot(data = estrogen, aes(x=group, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3') + facet_wrap(~protocol, ncol = 3)
#replicate groups are different because each replicate group was given a different dose of EE and ZM

ggplot(data = estrogen, aes(x=group, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3')
#Between all protocols the groups show similar trends of more EE leads to higher uterus weight. Interaction between EE and ZM may be significant

ggplot(data = estrogen, aes(x=protocol, y=log(uterus))) + geom_point(alpha = .5 , color = 'blue') + geom_smooth(method= 'lm' ,col= 'red3')
#Difference in weight, but expected since the age of the rats differed in the protocols

plot(log(uterus)~mweight, data = estrogen)
# very obvious clustering of mature and inmature rats
```


```{r modeling1, include = FALSE}
partial_estrogen_model1 <- lmer(log(uterus)~ mweight + EE + ZM + (1 | lab) + (1| protocol), data = estrogen)
summary(partial_estrogen_model1)
#ranef(partial_estrogen_model1)
#T values for these are large, and therefore the model is acceptable
car::vif(partial_estrogen_model1)
AIC(partial_estrogen_model1)
BIC(partial_estrogen_model1)

# weight is not very significant 
```

```{r modeling2, include = FALSE}

# created mature and replace weight to mature and adding level for protocol, qq plot not improving
two_level_estrogen_model <- lmer(log(uterus)~ EE + ZM + mature + (1| protocol) + (1 | lab), data = estrogen)
summary(two_level_estrogen_model)
car::vif(two_level_estrogen_model)
AIC(two_level_estrogen_model)
BIC(two_level_estrogen_model)

# tried to log weight
logx_estrogen_model <- lmer(log(uterus)~ log(mweight) + EE + ZM + (1 | protocol) + (1 | lab), data = estrogen)
summary(logx_estrogen_model)
AIC(logx_estrogen_model)
BIC(logx_estrogen_model)

# then for answering sensitivity of each protocol in detecting EE and ZM levels, we tried to add EE and ZM random slope effect
two_level_estrogen_model1 <- lmer(log(uterus)~ EE + ZM + mature + (1+EE+ZM| protocol) + (1 | lab), data = estrogen)
# converging issue here!!! NEED TO CHANGE
```


```{r modeling3, include = FALSE}
# because converge issue, add protocal as a normal predictor and added two interaction terms to see the sensitivity based on Protocol A as the baseline. Assuming baseline A has no sensitivity. 
full_estrogen_model <- lmer(log(uterus)~EE + ZM + protocol + (1 | lab) + EE:protocol + ZM:protocol, data = estrogen)


#adding group in because it is an indicator of the biological (maybe) interaction between EE and ZM that cannot be captured by including the interaction term of EE and ZM
summary(full_estrogen_model)
ranef(full_estrogen_model)
AIC(full_estrogen_model)
BIC(full_estrogen_model)
car::vif(full_estrogen_model)
qqnorm(residuals(full_estrogen_model));qqline(residuals(full_estrogen_model))
#Including protocol means weight now has a lower t value, therefore dropping weight for next model
```



```{r include = FALSE,echo = FALSE}

plot(full_estrogen_model)
#equal variance is violated, they are not scattered around 0
#Linearity is violated, there are downward trends in the second half of the graph

# resid_full <- resid(full_estrogen_model)
# plot(estrogen$mweight, resid_full)
# plot(estrogen$EE, resid_full)
# #We are underfitting for 1mg dose and overfitting for 10 mg dose
# plot(estrogen$ZM, resid_full)
# plot(estrogen$protocol, resid_full)
# plot(estrogen$group, resid_full)

qqnorm(residuals(full_estrogen_model));qqline(residuals(full_estrogen_model))
# not very good 


```

##### Model Selection Process

From EDA, we found that there are only 3 different doses for ZM and around 9 different doses for EE. We were thinking about making EE and ZM as categorical variables, but the problem of making them into categorical variable would result in loss of information since 1 mg is not just different from 10 mg, it is 1/10 of 10 milligrams. 

The first model we tried was:
$$\begin{aligned}
Log (Uterus)= & (B_0 + \gamma_{lab} + \eta_{protocal}  ) + B_1EE + B_2ZM + B_3MWeight + \varepsilon_{ijk} \\
\end{aligned}$$

The AIC for this model is 3314, and BIC is 3356. We expected both score to be smaller for our final model. Mean centered weight is not very significant with -0.72 t-value but the protocal variable does contain information of the weight of the rat. There isn't multicollinearity issue in the model by checking the VIF, but the normality assumption seem violated from the qq-plot. It seems that weight cannot distinguish mature vs inmature rats specified in the protocols. Instead of including the weight, we created another binary variable of mature status. Because protocal A&B are done on inmature rats, and C&D were done on mature rats, the mature variable can help identify this difference. The mature variable has a t-value of 40.33, it turns out to be very significant predictor. However the normality issue is still not solved. We keep the problem aside for now and tried to see sensitivity of different protocols detecting eefects ZM or EE by adding random slope effects on protocol. However, the model had a converge issue. We decided to put protocal as a normal predictor instead of a level predictor, and add interaction term of it with EE and ZM to see the sensivitiy with base case to be protocol A. We also replaced weight back in to the model in hoping of it distinguish the mature vs inmature rats. However, mean centered weight is not very significant in the model also it caused a multicollinearity issue between weight and protocol. We throw weight out in our final model. 


##### Model

Our final model is:
$$\begin{aligned}
Log (Uterus)= & (B_0 + \gamma_{lab} ) + B_1EE + B_2ZM  + B_3Protocal + B_4Protocol*EE + B_5Protocol*ZM + \varepsilon_{ijk} \\
\end{aligned}$$

We checked our model assumptions by plot residual vs fitted and qq-plot. The model still does not satisify the normality assumption. There are some kind of clustering in the residual plot, because we cannot distinguish mature vs inmature rats and also the variance are not very constant for the clusters.

```{r final_model, echo = FALSE, fig.width =3, fig.height =3, fig.align='center', fig.show='hold'}
qqnorm(residuals(full_estrogen_model));qqline(residuals(full_estrogen_model))
plot(full_estrogen_model)
```

There was another approach we tried which is adding the group as an extra predictor variable in the model and this issue will be fixed. However, when we added the group variable, we are losing information on EE and ZM and the interpretation would not make sense to say different groups of rats can have different impacts. We also checked the issue of multicollinearity, the VIF for each predictors in our final model are all between 2.6 to 5.6. This is the best model we have considering all the questions needed to be answered. The AIC score for this model is 3228 and BIC is 3310. 








